Paper Analysis

### 2024-07-25 Paper_WriteUp/Paper_Plots.qmd
many updates to the plots to reflect updates to the paper

@@

result |> 
  unnest_longer(session_length) |> 
  mutate(multiple = if_else(session_length > 10, TRUE, FALSE)) |>  
  summarise(sum(multiple)/n())

@@

```{r plot global parameters}
data_path <- here("results/lichess1700-1900")

@@

@@

## then the winner effects plot

winner_plot_17_19 <- player_pars |> 
  filter(param == 2) |> 
  ggplot(aes(value)) +
  geom_histogram() +
  facet_wrap(~player_name, scales = "free", ncol = 6) +
  theme_6col() +
  labs(x = element_blank(), y = element_blank()) +
  scale_x_continuous(breaks = breaks_pretty(n = 3))

@@

```{r}
data_path <- here("results/lichessGrandmasters")

@@

player_id <- as_tibble(bullet_gm) |> 
  mutate(id = row_number()) |> 
  rename(player_name = value)

@@

## then the winner effects plot

winner_plot_gm <- player_pars |> 
  filter(param == 2) |> 
  ggplot(aes(value)) +
  geom_histogram() +
  facet_wrap(~player_name, scales = "free", ncol = 5) +
  theme_5col() +
  labs(x = element_blank(), y = element_blank()) +
  scale_x_continuous(breaks = breaks_pretty(n = 3))

@@

Here there are only 21 rather than 25 players at the moment, this will be
updated currently.

@@

## global effects first

data_path <- here("results/lichessGrandmasters")

@@

@@

@@

@@

```{r}
data_path <- here("results/lichessGrandmasters")

@@

## this the global and winner effect plots

path_id <- 1
data_path <- res_data_path[path_id]

@@

## indiv winner effects
random_effect_first <- fit_17_19_first$draws() %>% as_draws_df() %>%
  select(starts_with("beta[")) %>%
  pivot_longer(cols = everything()) %>%
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) %>% 
  mutate(when = "Initial")

@@

winner_effect_plot <- betas %>% 
  left_join(player_id, by = "id") %>% 
  mutate(player = factor(player, levels = users)) %>% 
  filter(param == 2) %>% 
  mutate(id = factor(id, levels = c(1:10))) %>% 
  ggplot(aes(value, fill = when)) +
  geom_histogram() +
  geom_vline(xintercept = 0, alpha = 0.5) +
  scale_x_continuous(breaks = c(-0.25, 0.25)) +
  # xlim(c(-1.5, 1.5)) +
  facet_grid(rows = vars(when), cols = vars(player), scales = "free") +
  labs(fill = element_blank(), y = element_blank(), 
       # title = "Individual Winner Effects, Selected 1700-1900 Bullet Players",
       x = element_blank()) +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.text = element_text(size = axis_text_size - 4),
        axis.title = element_text(size = axis_title),
        legend.text = element_text(size = legend_text),
        strip.text = element_text(size = axis_title - 6),
        legend.position = "none",
        plot.title = element_text(size = title_size))

@@

```{r}
path_id <- 1
data_path <- res_data_path[path_id]

@@

p1 <- games_won %>% 
  mutate(focal_id = as.character(focal_id)) %>% 
  left_join(player_id, by = c("focal_id" = "id")) %>% 
  ggplot(aes(x = games_won)) +
  geom_histogram() +
  facet_wrap(~player, scales = "free", ncol = 5) +
  geom_vline(data = orig_games_won, 
             mapping = aes(xintercept = games_won), col = "red") +
  scale_x_continuous(breaks = breaks_pretty(n = 3)) +
  # labs(title = "Posterior Predictive Distribution, Initial Games",
  labs(y = element_blank(),
       x = "Number of Games Won") +
  theme_5col()

@@

```{r}
## time varying ppc for 1700-1900 players

draws <- fit_17_19_first$draws() %>% as_draws_df() %>%
  select(starts_with(c("beta[", "gamma"))) %>%
  mutate(draw = row_number()) %>% 
  pivot_longer(cols = "beta[1,1]":"beta[2,10]") %>% 
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) %>% 
  select(-name) %>% 
  pivot_wider(names_from = param, values_from = value) %>% 
  rename(beta1 = `1`, beta2 = `2`)

@@

last_ppc <- ppc_data %>%  
  left_join(player_id, by = "id") %>% 
  ggplot(aes(x = games_won)) + 
  geom_histogram() + 
  geom_vline(data = future_results, 
             mapping = aes(xintercept = total), col = "red") +
  facet_wrap(~player, scales = "free_x", ncol = 5) +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
  labs(y = element_blank(), x = "Number of Games Won") +
  scale_x_continuous(breaks = breaks_pretty(n = 3)) +
  theme_5col()

@@

This data is complete, just need to create the plots here.
Here we use the 20 players with the most games only.

## these don't look right just now

perm_data_path <- here("results/lichess1700-1900/perm")

path_id <- 1
data_path <- res_data_path[path_id]
## read all the files in this directory
(perm_files <- list.files(perm_data_path))

@@

ggsave(plot = perm_17_19_global, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "perm_17_19_global.png")), dpi = 600,
       height = 5, width = 7)

## get the selected users for this, top 20 players

@@

ggsave(plot = perm_17_19_winner, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "perm_17_19_winner.png")), dpi = 600,
       height = 5, width = 7)

@@

```{r}
## just do ppc plots for now

path_id <- 4
data_path <- res_data_path[path_id]

fit_gm_first <- readRDS(file =  here(data_path, "time",
                                        "select_users_first_bullet.RDS"))
fit_gm_last <- readRDS(file =  here(data_path, "time",
                                        "select_users_last_bullet.RDS"))

@@

p1 <- games_won %>% 
  mutate(focal_id = as.character(focal_id)) %>% 
  left_join(player_id, by = c("focal_id" = "id")) %>% 
  ggplot(aes(x = games_won)) +
  geom_histogram() +
  facet_wrap(~player, scales = "free", ncol = 5) +
  geom_vline(data = orig_games_won, 
             mapping = aes(xintercept = games_won), col = "red") +
  scale_x_continuous(breaks = breaks_pretty(n = 3)) +
  # labs(title = "Posterior Predictive Distribution, Initial Games",
  labs(y = element_blank(),
       x = "Number of Games Won") +
  theme_5col()

@@

last_ppc <- ppc_data %>%  
  left_join(player_id, by = "id") %>% 
  ggplot(aes(x = games_won)) + 
  geom_histogram() + 
  geom_vline(data = future_results, 
             mapping = aes(xintercept = total), col = "red") +
  facet_wrap(~player, scales = "free_x", ncol = 5) +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
  labs(y = element_blank(), x = "Number of Games Won") +
  scale_x_continuous(breaks = breaks_pretty(n = 3)) +
  theme(axis.text = element_text(size = axis_text_size - 2),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.spacing.x = unit(0.5, "lines"),
        plot.title = element_text(size = title_size),
        axis.title = element_text(size = axis_title),
        legend.text = element_text(size = legend_text),
        strip.text = element_text(size = axis_title - 6))

@@

ggsave(plot = last_ppc, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "gm_ppc_2.png")), dpi = 600,
       height = 5, width = 7)

@@

perm_gm_winner <- player_pars |> 
  filter(param == 2) |> 
  ggplot(aes(mean)) +
  geom_histogram() +
  facet_wrap(~player_name, ncol = 5, scales = "free") +
  labs(x = element_blank(), y = element_blank()) +
  geom_vline(data = true_beta |> filter(param == 2), aes(xintercept = mean),
             col = "red") +
  theme_5col()

@@

perm_gm_player <- player_pars |> 
  filter(param == 1) |> 
  ggplot(aes(mean)) +
  geom_histogram() +
  facet_wrap(~player_name, ncol = 5, scales = "free") +
  labs(x = element_blank(), y = element_blank()) +
  geom_vline(data = true_beta |> filter(param == 2), aes(xintercept = mean),
             col = "red") +
  theme_5col()

@@

## should make this 12 players so it looks a bit better

@@

>>>>

## with more than 10 games
result |> 
  unnest_longer(session_length) |> 
  mutate(multiple = if_else(session_length > 10, TRUE, FALSE)) |>  
  summarise(sum(multiple)/n())

## repeat for blitz games

outcomes_blitz <- blitz_data |> 
  # group_by(Username) |> 
  select(Username, UTCDate, UTCTime) |> 
  mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
  group_by(Username) |> 
  arrange(date_time) |> 
  mutate(gap = date_time - lag(date_time),
         gap_seconds = as.numeric(gap)) |> 
  mutate(less_7 = if_else(gap_seconds <= 420, TRUE, FALSE, missing = FALSE))

result_blitz <- outcomes_blitz |> 
  summarise(session_length = list(get_true_sequences(less_7)))

result_blitz |> 
  unnest_longer(session_length) |> 
  mutate(multiple = if_else(session_length > 1, TRUE, FALSE)) |>  
  summarise(sum(multiple)/n())

@@

## proportion of sessions with more than 1 game
result_gm |> 
  unnest_longer(session_length) |> 
  mutate(multiple = if_else(session_length > 1, TRUE, FALSE)) |>  
  summarise(sum(multiple)/n())

result_gm |> 
  unnest_longer(session_length) |> 
  mutate(multiple = if_else(session_length > 10, TRUE, FALSE)) |>  
  summarise(sum(multiple)/n())

@@

```{r plot global parameters for 17-19}
data_path <- here("results/lichess1700-1900")

@@

```{r player level parameters for 17-19}

@@

## see what the smallest and largest posterior mean is 
## for the winner effects
player_pars |> 
  group_by(player_name, param) |> 
  summarise(post_mean = mean(value)) |> 
  filter(param == 2) |> 
  ungroup() |> 
  slice_max(order_by = post_mean, n = 5)

player_pars |> 
  group_by(player_name, param) |> 
  summarise(post_mean = mean(value)) |> 
  filter(param == 2) |> 
  ungroup() |> 
  slice_min(order_by = post_mean, n = 5)

## see how many contain 0
player_pars |> 
  group_by(player_name, param) |> 
  filter(param == 2) |> 
  summarise(lower = quantile(value, prob = 0.025),
            upper = quantile(value, prob = 0.975)) |> 
  mutate(contain_zero = ifelse(upper > 0 & lower < 0,
                         1, 0)) |> 
  filter(contain_zero == 0)

## then the winner effects plot
winner_plot_17_19 <- player_pars |> 
  filter(param == 2) |> 
  ggplot(aes(value)) +
  geom_histogram() +
  facet_wrap(~player_name, scales = "free", ncol = 6) +
  theme_6col() +
  labs(x = element_blank(), y = element_blank()) +
  scale_x_continuous(breaks = breaks_pretty(n = 3))

@@

```{r global parameters for gm data}
data_path <- here("results/lichessGrandmasters")

@@

```{r player level parameters for gm data}
player_id <- as_tibble(bullet_gm) |> 
  mutate(id = row_number()) |> 
  rename(player_name = value)

@@

## see what the smallest and largest posterior mean is 
## for the winner effects
player_pars |> 
  group_by(player_name, param) |> 
  summarise(post_mean = mean(value)) |> 
  filter(param == 2) |> 
  ungroup() |> 
  slice_max(order_by = post_mean, n = 5)

player_pars |> 
  group_by(player_name, param) |> 
  summarise(post_mean = mean(value)) |> 
  filter(param == 2) |> 
  ungroup() |> 
  slice_min(order_by = post_mean, n = 5)

## count how many intervals contain 0
## see how many contain 0
player_pars |> 
  group_by(player_name, param) |> 
  filter(param == 2) |> 
  summarise(lower = quantile(value, prob = 0.025),
            upper = quantile(value, prob = 0.975)) |> 
  mutate(contain_zero = ifelse(upper > 0 & lower < 0,
                         1, 0)) |> 
  filter(contain_zero == 0)

## then the winner effects plot
winner_plot_gm <- player_pars |> 
  filter(param == 2) |> 
  ggplot(aes(value)) +
  geom_histogram() +
  facet_wrap(~player_name, scales = "free", ncol = 5) +
  theme_5col() +
  labs(x = element_blank(), y = element_blank()) +
  scale_x_continuous(breaks = breaks_pretty(n = 3))

@@

```{r interpret changes in win probability for gm}
## interpret the individual winner effects
## assume wins 50% of all games

1/(1 + exp(-1 * (0)))

## if won previous
1/(1 + exp(-1 * (0 + 0.216 * 0.5)))

## if lost previous
1/(1 + exp(-1 * (0 + 0.216 * - 0.5)))

## what about if win probability overall was 40%
1/(1 + exp(-1 * (0 + 0.216 * 0.6)))

## 35%, pretty low
1/(1 + exp(-1 * (0 + 0.216 * 0.65)))

@@

```{r compare global params n-1 vs n-10}
## global effects first
data_path <- here("results/lichessGrandmasters")

@@

## check the proportion of 95% intervals which contain zero
bind_rows(player_pars_pre, player_pars_pre_10) |> 
  filter(param == 2) |> 
  group_by(player_name, model) |> 
  summarise(lower = quantile(value, prob = 0.025),
            upper = quantile(value, prob = 0.975)) |> 
  mutate(contain_zero = ifelse(upper > 0 & lower < 0,
                         1, 0)) |> 
  group_by(model) |> 
  summarise(prop = sum(contain_zero)/ n())

## do the effects change in direction
bind_rows(player_pars_pre, player_pars_pre_10) |> 
  filter(param == 2) |> 
  group_by(player_name, model) |> 
  summarise(lower = quantile(value, prob = 0.025),
            upper = quantile(value, prob = 0.975)) |> 
  mutate(contain_zero = ifelse(upper > 0 & lower < 0,
                         1, 0)) |>
  print(n = 50)

@@

<!-- down to here in terms of updating the code, making sure it's complete -->

@@

```{r bullet v blitz for 17-19}

@@

```{r bullet v blitz gm}
data_path <- here("results/lichessGrandmasters")

@@

```{r ppc 17-19}
## this the global and winner effect plots
path_id <- 1
data_path <- res_data_path[path_id]

@@

users <- readRDS(file = here(data_path, "time", "select_users_bullet.RDS"))
## indiv winner effects
random_effect_first <- fit_17_19_first$draws() %>% as_draws_df() %>%
  select(starts_with("beta[")) %>%
  pivot_longer(cols = everything()) %>%
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) %>% 
  mutate(when = "Initial")

@@

winner_effect_plot <- betas %>% 
  left_join(player_id, by = "id") %>% 
  mutate(player = factor(player, levels = users)) %>% 
  filter(param == 2) %>% 
  mutate(id = factor(id, levels = c(1:10))) %>% 
  ggplot(aes(value, fill = when)) +
  geom_histogram() +
  geom_vline(xintercept = 0, alpha = 0.5) +
  scale_x_continuous(breaks = c(-0.25, 0.25)) +
  # xlim(c(-1.5, 1.5)) +
  facet_grid(rows = vars(when), cols = vars(player), scales = "free_x") +
  labs(fill = element_blank(), y = element_blank(), 
       # title = "Individual Winner Effects, Selected 1700-1900 Bullet Players",
       x = element_blank()) +
  theme_10col()
  # theme(axis.ticks.y = element_blank(),
  #       axis.text.y = element_blank(),
  #       axis.text = element_text(size = axis_text_size - 4),
  #       axis.title = element_text(size = axis_title),
  #       legend.text = element_text(size = legend_text),
  #       strip.text = element_text(size = axis_title - 6),
  #       legend.position = "none",
  #       plot.title = element_text(size = title_size))

@@

```{r ppc1 for 17-19}
path_id <- 1
data_path <- res_data_path[path_id]

@@

p1 <- games_won %>% 
  mutate(focal_id = as.character(focal_id)) %>% 
  left_join(player_id, by = c("focal_id" = "id")) %>% 
  ggplot(aes(x = games_won)) +
  geom_histogram() +
  facet_wrap(~player, scales = "free", ncol = 5) +
  geom_vline(data = orig_games_won, 
             mapping = aes(xintercept = games_won), col = "red") +
  scale_x_continuous(breaks = breaks_pretty(n = 3)) +
  # labs(title = "Posterior Predictive Distribution, Initial Games",
  labs(y = element_blank(),
       x = "Number of Games Won") +
  theme_5col() +
  theme(axis.title.x = element_text(size = 10,
                                    margin = margin(t = 10, b = 10)))

@@

```{r ppc2 for 17-19}
## time varying ppc for 1700-1900 players
draws <- fit_17_19_first$draws() %>% as_draws_df() %>%
  select(starts_with(c("beta[", "gamma"))) %>%
  mutate(draw = row_number()) %>% 
  pivot_longer(cols = "beta[1,1]":"beta[2,10]") %>% 
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) %>% 
  select(-name) %>% 
  pivot_wider(names_from = param, values_from = value) %>% 
  rename(beta1 = `1`, beta2 = `2`)

@@

last_ppc <- ppc_data %>%  
  left_join(player_id, by = "id") %>% 
  ggplot(aes(x = games_won)) + 
  geom_histogram() + 
  geom_vline(data = future_results, 
             mapping = aes(xintercept = total), col = "red") +
  facet_wrap(~player, scales = "free_x", ncol = 5) +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
  labs(y = element_blank(), x = "Number of Games Won") +
  scale_x_continuous(breaks = breaks_pretty(n = 3)) +
  theme_5col() +
  theme(axis.title.x = element_text(size = 10,
                                    margin = margin(t = 10, 
                                                    b = 10,
                                                    unit = "pt")))

@@

```{r perm for 17-19}
perm_data_path <- here("results/lichess1700-1900/perm")

path_id <- 1
data_path <- res_data_path[path_id]
(perm_files <- list.files(perm_data_path))

@@

ggsave(plot = perm_17_19_global, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "perm_17_19_global.png")), dpi = 600,
       height = 5, width = 7)
```

```{r perm winner effects}
## get the selected users for this, top 20 players
path_id <- 1  ## for 1700-1900
data_path <- all_data_path[path_id]
files <- list.files(data_path)

@@

ggsave(plot = perm_17_19_winner, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "perm_17_19_winner.png")), dpi = 600,
       height = 5, width = 7)
```

@@

```{r ppc1 and ppc2 for gm}
## just do ppc plots for now
path_id <- 4
data_path <- res_data_path[path_id]
fit_gm_first <- readRDS(file =  here(data_path, "time",
                                        "select_users_first_bullet.RDS"))
fit_gm_last <- readRDS(file =  here(data_path, "time",
                                        "select_users_last_bullet.RDS"))

@@

p1 <- games_won %>% 
  mutate(focal_id = as.character(focal_id)) %>% 
  left_join(player_id, by = c("focal_id" = "id")) %>% 
  ggplot(aes(x = games_won)) +
  geom_histogram() +
  facet_wrap(~player, scales = "free", ncol = 5) +
  geom_vline(data = orig_games_won, 
             mapping = aes(xintercept = games_won), col = "red") +
  scale_x_continuous(breaks = breaks_pretty(n = 3)) +
  # labs(title = "Posterior Predictive Distribution, Initial Games",
  labs(y = element_blank(),
       x = "Number of Games Won") +
  theme_5col() +
  theme(axis.title.x = element_text(size = 10,
                                   margin = margin(t = 10, b = 10, unit = "pt")),
        panel.spacing.x = unit(0.7, "lines"))

@@

last_ppc <- ppc_data %>%  
  left_join(player_id, by = "id") %>% 
  ggplot(aes(x = games_won)) + 
  geom_histogram() + 
  geom_vline(data = future_results, 
             mapping = aes(xintercept = total), col = "red") +
  facet_wrap(~player, scales = "free_x", ncol = 5) +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
  labs(y = element_blank(), x = "Number of Games Won") +
  scale_x_continuous(breaks = breaks_pretty(n = 3)) +
  theme_5col() +
  theme(axis.title.x = element_text(size = 10,
                                   margin = margin(t = 10,
                                                   b = 10,
                                                   unit = "pt")))

@@

ggsave(plot = last_ppc, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "gm_ppc_2.png")), dpi = 600,
       height = 5, width = 7)
```

@@

perm_gm_winner <- player_pars |> 
  filter(param == 2) |> 
  ggplot(aes(mean)) +
  geom_histogram() +
  facet_wrap(~player_name, ncol = 5, scales = "free") +
  labs(x = element_blank(), y = element_blank()) +
  geom_vline(data = true_beta |> filter(param == 2), aes(xintercept = mean),
             col = "red") +
  theme_5col() +
  scale_x_continuous(breaks = breaks_pretty(n = 3))

@@

perm_gm_player <- player_pars |> 
  filter(param == 1) |> 
  ggplot(aes(mean)) +
  geom_histogram() +
  facet_wrap(~player_name, ncol = 5, scales = "free") +
  labs(x = element_blank(), y = element_blank()) +
  geom_vline(data = true_beta |> filter(param == 2), aes(xintercept = mean),
             col = "red") +
  theme_5col() +
  scale_x_continuous(breaks = breaks_pretty(n = 3))

@@

## should we make this 12 players so it looks a bit better?
```

@@

## Model Checking for the Other Cohorts
